\chapter{Who This Manual is For}

I sincerely hope this manual is extremely useful to you as you attempt to solve one of the most common problems in individually owned technical services firms: the lack of proper positioning.  If at any point in working your way through this manual you run across problems, I encourage you to contact me directly. I'm Philip Morgan, and I help custom software development shops get more leads without hiring a sales person. I use positioning, education-based content marketing, and marketing automation to make that happen. You can reach me via email at \href{mailto:philip@philipmorganconsulting.com}{philip@philipmorganconsulting.com}.

\section{Who This Manual is For}

This manual will be most helpful to a narrow range of people:

\begin{itemize}
\item Owners of technical services firms like programming consultancies, web or digital agencies, and SaaS companies.
\item Independent consultants or freelancers of many stripes, though all my examples in this manual are specific to the technical rather than the creative end of the professional services spectrum.
\item People who are considering making the leap to solo self-employment but haven't done it yet.
\item Self-employed people who need to be making more money within 12 months, but can invest 6 months of effort to lay the foundation.
\end{itemize}

This manual will \emph{not} be immediately helpful to you if you have very limited work experience or ``the wolf is at the door'' in terms of your finances. Developing a desirable market position requires discipline in client selection and negotiation that will be very difficult to muster if you have little savings, credit, or leeway in terms of saying no to work in the short term. Making a good positioning choice requires a certain amount of self-awareness about your strengths and preferences, and you just can't discover that stuff without a year or two of full-time client work under your belt.

Finally, this manual is meant to be concise. How many 80,000-word business books have you read that should have been 20,000 words? So many business books are one meaningful but straightforward insight padded out with redundant content to make a publisher happy. This one is not.

Even with a few interruptions you should be able to read this book cover to cover in 2 hours or less. I’m confident you’ll get a \emph{world class} ROI on that modest time investment.